Search for Spatially Extended Fermi-LAT Sources Using Two Years of Flight Data

### 2011-06-22 extended_catalog/paper/extended_catalog.tex
Imporved draft + plots

% TODO:
%   Add contours to images.

@@

\begin{abstract}
We present a new method for analyzing spatially extended sources with
the Fermi Large Area Telescope (LAT), the primary science instrument
on the {\em Fermi Gamma-ray Space Telescope (Fermi)}.  We provide a
series of Monte Carlo studies that were done to validate this tool.
We then calculate the Fermi LAT's sensitivity to spatially exnteded
sources.  We then apply this tool to test all sources in the 2FGL Two
Year Catalog for extension.\cite{2fgl} We then report on the discovery
of XXXXXXXXXXXXXX new extended sources.
\end{abstract}

@@

The Large Area Telescope (LAT) is a pair conversion telescope on the
Fermi Gamma-Ray Space Telescope. It has been surveying the Gamma-ray
sky since June 2008.  Fermi is particularly well suited as a survey
instrument because of its to survey the broad energy coverage ($20\mev$
to $>300\gev$), a wide field of view ($\sim 2.4 \sr$) and large effective
area ($\sim 8000 \cm^2$ at $>1\gev$).

@@

Being able to spatially resolve a fermi source is important for several
reasons. Most importantly, the spatial morphology can be used to help
identify a Fermi source. Because of the large point spread function (PSF),
for any source there are often several counterparts which it could be
associated with. Finding a matching morphology can uniquly identify the
source emitting the Gamma-rays.

@@

Nevertheless, morphological studies of sources in the $\gev$ energy range using
the Fermi LAT is challenging because of the significantly energy
dependent PSF and because of systematic errors in the
galactic diffuse emission. The Fermi PSF varies from over $XXXXXXXXXXXXXXX\deg$
at $100\mev$ to $XXXXXXXXXXXXXXXX\deg$ at $100\gev$ so the higher energy photons 
are significantly more important for resolving a source. On the other hand, 
we tend to run out of photons at higher energies so a detailed statistical 
test must be used to maximize our sensitivity.

@@

%   * Difficulty of studying spatially extended sources
%   * Pointlike, great features
%   * modifying it to fit spatially extended sources.
%      * semi-analytic convolution formula
%   * importance of fitting position + extension + spectrum simultaneously to apply statistical significance
%   * Monte Carlo Validation
%

Performing morphological studies of sources in the $\gev$ energy range using
the Fermi LAT is challenging primarily because of the significantly energy
dependent PSF and because of systematic errors in the
galactic diffuse emission. The Fermi PSF varies from over $XXXXXXXXXXXXXXXX\deg$
at $100\mev$ to $XXXXXXXXXXXXXXXX\deg$ at $100\gev$ and so the higher energy photons are
significantly more important for studying the shape of sources. One could
ideally look at just the highest energy photons, but the trade off is that
for typical sources many more source events are expected at low energy.
For this reason, it is especially important to use as much of the data
as possible.

@@

\pointlike additionaly bins the Fermi Gamma-rays by their conversion
type, which corresponds to whether the photon converted in the front or
back part of the detector. Since the PSF is smaller for front entering
events, one would expect binning each conversion type independently to
improve our sensitivity.

@@

Then, the extended source's position error is calculated using the
same code employed to calculate position errors of point sources. This
code is used to calculate the localization error of all point sources
kn the catalog\cite{2fgl}. It works by fitting an ellipse to the
likelihood surface near the maximum and using this fit ellipse to estimate
an elliptical localization errors. By reporting two localization errors
as well as a localization error angle, this method explicitly takes care
of covariance between the two position coordinates.

% Make an extension profile plot.
% show it in figures an ddescirbeit.

@@

This is demonstrated in figure~\ref{compare_disk_gauss}.  The plot shows
(in black) the PSF that would be observed for a source of spectral index
2. The dashed red (blue) line shows the probabability density function
for a Gaussian (uniform disk) of width $0.5\deg$.  The solid red (blue)
line shows the probability density function for the same Gaussian (uniform
disk) convolved with the PSF of an index 2 source.  
figure~\ref{compare_disk_gauss} shows the analysis using an energy range
of 1\gev to 100\gev whereas the right plot shows an analysis using an
energy range from 10\gev to 100\gev.

@@

Finally, we note that comparing observed morphologies to 
counts maps can be misleading.  This effect can best be seen
in figure~\ref{compare_r68} where we have simulted a
extended source with 
a spatial model that is a uniform disk of extension 0.5\deg,
integral flux between 1\gev and 100\gev of $3e-8\ph/\cm^2/s$,
and a spectral index of 2.

@@

@@

The half $\chi^2_1$ distribution was found by a monte carlo study
to hold for the special case of source finding of EGRET sources using the
maximum likelihood technique.  They were comparing the hypothesis of a
source to the hypothesis of just a background\cite{mattox_egret}.
Their argument for why this null hypothesis distribution was reasonable
was that half of the time, statistical underfluctuations would cause
there to be fewer fewer photons then expected where the sources is being
tested. Since the maximim likelihood machinery cannot fit negative
sources (which would lead to negative model predicted for which the
poisson likelihood is undefined), half of the tests would not have the
likelihood improved by adding a source to the model and $\ts=0$.
It is for the case that the null distribution has

@@

We performed a dedicated monte carlo study to validate \pointlike for
studying spatially extended sources and also to demonstrate that we can
use $\tsext$ as a test of the statistical significance
of a source's extension.  The monte carlo study involved simulating point
sources of varying spectral models on top of a simulated background. The
simulated point source was then fit using \pointlike as both an extended
source and a point source and $\tsext$ was calculated.

Fermi LAT data was simulated using program \gtobssim which is
distributed publically by the Fermi Science Support Center
(\cite{fssc}).  The sources were simulated with a powerlaw spectral
model with 6 Were given 6 $>100\mev$ fluxes ranging from $3\times 10^{-9}
\ph/\cm^2/s$ to $10^{-6} \ph/\cm^2/s$.  The sources were then simulated
with spectral index ($\gamma$) values of 1.5, 2, 2.5, and 3.  These
values were picked to probe a representative range measurable values.
For particular spectral indices, only the fluxes which would create a
statistically significant source detection ($\ts>25$) where selected,
since we are only interested in testing for extension sources with a
significant extent. The sources were simulated on top of an isotropic
powerlaw background with a Sreekumar-like spectrum ($>100\mev$ flux
of 1.5e-5 and spectral index 2.1) (\cite{sreekumar_isotropic}).

The monte carlo simulation was perforormed over a one year time interval using
a default rocking profile and an assumed a represntative
livetime fraction of 0.8.
The reconstruction was perfromed with events from an energy range of
$1\gev$ to $100\gev$ using 4 energy bins per decade and using the Pass 7 Version 6
source instrument response function (IRFs) to be consistent with the following analysis.

@@

The results of the monte carlo simulation are shown in
figure~\ref{ts_ext_mc}.  The cumulative density function (CDF) is plotted
against $\tsext$ For reference, the CDF for $\chi^2_1/2$,
suggested by Wilks' Theorem, is also plotted for comparison. Overall,
there is good agreement between Wilks' theorem, although the agreement is
not perfect.  It should be noted that the discrepancy seems to be worst
for bright sources which seems to imply that numerical errors in the
way \pointlike models the PSF and performs the convolution become more
apparent for sufficiently bright sources.  Other possible reasons for
departure from Wilks' theorem might include \pointlike ignoring energy
dispersion. Energy dispersion will change the shape as a function of
energy. \pointlike may also run into possibly problems with the fitting
algorithm getting stuck in false minimum. It is important to emphasize
that if
our emperical curve lies to the left of the curve we assume to estimate
significance, we would underestimate the statistical significance of
our result. Since this is the case for most of our simluations, we are
reasonable confident that $\sqrt{\tsext}$ can be safely
used as a measure of the statistical significance of a source's extension.

@@

We performed another monte carlo study to determine the Fermi LAT's
sensitivity to spatially extended sources. The sensitivity to extension
is defined as the flux at which 
the average value of $\tsext$ for a source is 
e equal to 25 (for a given spectral model). This correseoncs
to an average extension significance of $5\sigma$.  This is not the
sensitivity to detect an extended source but instead the sensitivity to
resolve an already detected source as being extended.

@@

To calculate the LAT's sensitivity to spatially extended sources,
we simulated sources with a uniform radially symmetric
disk morphology on top of a powerlaw Sreekumar-like isotropic
spectrum (\cite{sreekumar_isotropic}).  The monte carlo
simulation was perforormed over a one year time interval using a default
rocking profile and an assumed a represntative livetime fraction of 0.8.
For a given extension and spectral index, we calculated the
sensitivity by varying the flux of a source until we found 
when $\tsext=25$. In particular, we hand picked a flux range
where $\tsext<25$ and $\tsext>25$ and simulated 10 points with
a log uniform distribution in this range and performed an extension
test for each of these fluxes. Since the measured signifciance is
a statistical which will vary from realization to realization,
we then repeated this procedure $>15$ times for each spectral
and spatial model.
By fitting a line to the flux vs $\tsext$ data, we were able to
estimage the flux at which $\tsext=25$.

Our first result is the extension sensitivity for sources of varying
spectral indices. The result is shown in figure~\ref{index_sensitivity}.
We computed the sensitivty for sources with a spectral index of 1.5,
2, 2.5, and 3 and for sources of 20 extensions varying from $0.1\deg$
to $2.0\deg$. The $y$ axis is the $100\mev$ to $100\gev$ flux of the
source and the $x$ axis is the monte carlo extension. The curves then
represent the division in this extension and flux space where one gets
on average $\tsext=25$.
The solid lines of the plot are the sensitivity computed when using
with all events from $100\mev$ to $100\gev$. The dashed lines is the
same sensitivity but only using events from from $1\gev$ to $100\gev$.

This results shows that our extension sensitivity is
significantly better for harder sources where there are more high
energy events. Furthermore, our sensitivity is best for sources with an
extension around $0.5\deg$.  our sensitivity gets significantly worse
for very small sources and gets slightly worse for very large sources.

Futheremore, we note that except for the hardest or largest sources,
our sensitivity is almost entirly whether or not we use events from
$100\mev$ to $1\gev$. On the other hand, other systematic errors become
increasingly important at low energy so for our extension search we will
not use the events in this energy range.

We performed a second computation to see how the LAT's extension
sensitivity varied with increasing background, as is at lower latitude
in the galactic plane. To do so, we computed the sensitivity using the
extragalactic barground as well as 10 and 100 times the extragalactic
background. The latter result approximatly represents the intensity of
the galactic diffuse emission near the galactic center. The results are
shown in figure~\ref{diff_factor_sensitivity}. Two plots are presented,
one fitting only in the energy range between $1\gev$ and $100\gev$ and
one fitting only in the energy range between $10\gev$ and $100\gev$. For
each plot, the $y$ axis quotes the flux in the fit energy range and the
simulation was done only for a representivite spectral index 2 source.

@@

The LAT two year catalog presented a search for all
statistically significant sources that are detected in the $\gev$
sky\cite{2fgl}.  The catalog group's source finding method assumes
that the emission originated as point sources and what was presented was
a list of all sources fit assuming a point source spatial hypothesis. 
Published extended sources were modeled by the catalog group as 
being extended, but no attempt was made to test if other sources in the catalog
could be spatially resolved.

@@

As was shown, we gain very little in sensitivity using events with energy
below $1\gev$. On the other hand, the large PSF at low energy makes us
more succeptable to systmetic errors arising from source confusion due
to multiple point soruces and succeptable to incorrect modeling of the
galactic diffuse emission.  For that reason, we performed our search
using events only between $1\gev$ and $100\gev$.

We also performed a search for extended sources using events only between
10\gev and 100\gev. Even though we are testing the same sources, this
approach is complimentary because there are regions in the galactic
plane where at lower energies the diffuse emission dominates and is
likely not perfectly modeled. On the other hand, at the highest energies
accessible by the LAT, the galactic diffuse emission will be smaller and
we might expect to detect significantly harder sources. This approach is
also beneficial because harder sources with nearby neighbors that are
softer may end up being incorrectly fit at lower energies but correctly
found looking only at the high energy.  This is also beneficial in the
galactic plane looking for possible extended sources near pulsars which
are very bright at lower energy but cutoff before 10\gev.  This approach
of using only events above $10\gev$ was succesfully used to detecting
HESS J1825-137 with the Fermi LAT~\cite{fermi_hess_j1825}.

For each catalog source, we modified the source to have a uniform radially
symmetric disk spatial shape and used \pointlike to simultaneously fit
the position and extension of the source.  We then selected the sources
for which \pointlike found $\tsext>16$ and refit their spectral parameters
using \gtlike.

% XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
% NOTE ABOUT PRUNE aridus

@@

We expect most extended sources to be galactic sources since extragalactic
sources are typically too far away to be resolvable. Unfortunately,
the \gev emission in the galactic plane is dominated by anisotropic
galactic diffuse emission.  Modeling the galactic diffuse emission is
very difficult~\ref{first_diffuse_paper}  and finding sources on top of
this emission is notoriously difficult. This issue has been discussed
in the Fermi catalog~\ref{first_cat,2fgl}. Furthermore, the galactic plane
is very crowded and there is a degeneracy between an extended source
and multiple point sources. So it is often difficult to ensure you have
a good model for the region around a source .  Finding a source with
$\tsext>16$ is not sufficient for detecting an extended and we had to
perform many crosschecks to make sure we had a good fit to our region.

% XXXXXXXXXXXXXXXXXXXXX
% LIST OF All extended source candidates and how many of them are clearly bad.
% best we can do is sort through them.

% XXXXXXXXXXXXXXXXXXXXXXXXXXX
% Put example of poorly fit region and how it is difficult to know what
% is going on.
Example Bad fit~\ref{example_bad_fit}

For each of the extended source candidates where $\tsext>16$,
we performed several additional analyses to determine whether
our model for the region was believable and that we could therefore
find out source to be extended.
We did so for all sources found either at events above 1\gev or above 10\gev

First, we took the spatial and spectral model found using \pointlike
and performed a crosscheck by refitting the spectral parameters using 
\gtlike.
We used a square ROI $14.1\deg$ in diamater with $1/16\deg$ pixels
and the same energy binning.
% XXXXXXXXXXXXXX put note about how this binning was necessary for morphological analysis and how any smaller is no good.
% using gtlike. But how pointlike handles this automatically
The \gtlike analysis provided a second estimate of the spectral
parameters of the extended source as well as a second calculation of
$\ts_\text{disk}$, $\ts_\text{point}$, and 
$\tsext$. We only looked for extended sources which 
both \pointlike and \gtlike considered significantly extended.

% XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
% Note about how it is helfpul to look at high energy
% to avoid source confusion and even though a source
% might show up above 1GeV, it might be hard to realize
% wheter it is source confusion/diffuse emission, etc.
% but at high energy, less diffuse emission + better PSF
% = easier to find extended sources.

% Also note about how softer other sources nearby 
% might go away at higher energy

% XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
% NOTE about pointlike TS_ext generally being larger then
% gtlike TS_ext and why this may be (binning front/back evnets
% seperatly, using too large an energy bin, etc).
% 
% This is no longer true with better binning 1/16 instead of 1/8 degree binning.

Additionaly, we generated for each 
extended source candidate residual test statistic maps. 
The residual TS Map

Large
TS values in test

% XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
% Note about dual localization

We generated a residual test statistic maps to look for
, which is a map of the test
statistic adding new points
Figure~\ref{res_tsmaps} shows an example residual TS map.

% XXXXXXXXXXXXXXX add now about using powerlaw models for our source +
% setting e0 in geometric range of energy bin.

% XXXXXXXXXXXXXXXXXXXXX
% Discussion of how 
% counts slices are best for
% looking at overall diffuse emission.

% XXXXXXXXXXXXXX discussion of complicated regions having to simulatenously refit
% backgorund sources. Relocalizing sources.
% bg sources are obtained by catalog w/ > 100MeV analysis may
% not be idea for >1GeV & >10GeV analysis. Also, may require moving b/c
% extension will effect their extension.

@@

We validated our method by testing for extension all of the sources
which the Fermi Collaboration had previously published on.  We applied
no special method to test these sources but used the same pipeline which
was used to test all other catalog sources.  The results are published
in table~\ref{known_extended_sources}.  Using events with energies
greater than $1\gev$, our analysis succesfully fit the extension of the
the two galaxies the Large Magalengic Cloud and the Small Magalenic
Cloud.~\cite{SMC & LMC Paper}.  Our tool also successfully detected the
extension of the supernova remenants IC443, W28, W30, W44, and W51C.

Using events above $10\gev$, we were able to succesffuly fit the extesion
of HESS J1825-137\cite{fermi_hess_j1825}. Our analysis of MSH 15-52 using
events above $10\gev$ found an extension consistent with the publisehed
result, but with $\tsext=11.2$.\cite{msh1552}. This
is most likely because the published analysis used events above $6\gev$.

Our analysis failed to detect the extension of Vela X. Vela X can only be
successfully analyzed after applying a phase cut to removed the heavily
dominant emission from the pular.

>>>>

% TODO:
%   Add contours to images.
%   What other citations to add?

@@

\begin{abstract}
We present a new method for analyzing spatially extended sources with
the Fermi Large Area Telescope (LAT), the primary science instrument
on the {\em Fermi Gamma-ray Space Telescope (Fermi)}.  We provide a
series of Monte Carlo studies that were done to validate this tool.
We then calculate the Fermi LAT's sensitivity to spatially exnteded
sources.  We then apply this tool to test all sources in the 2FGL Two
Year Catalog for extension.\cite{second_cat} We then report on the discovery
of XXXXXXXXXXXXXX new extended sources.
\end{abstract}

@@

The Large Area Telescope (LAT) is a pair conversion telescope on the
Fermi Gamma-Ray Space Telescope. It has been surveying the $\gamma$-ray
sky since June 2008.  Fermi is particularly well suited as a survey
instrument because of its to survey the broad energy coverage (20 \mev
to >300 \gev), a wide field of view ($\sim 2.4 \sr$) and large effective
area ($\sim 8000 \cm^2$ at $>1 \gev$).

@@

Being able to spatially resolve a fermi source is important for several
reasons. Most importantly, the spatial morphology can be used to help
identify a Fermi source. Because of the large point spread function (PSF),
for any source there are often several counterparts which it could be
associated with. Finding a matching morphology can uniquly identify the
source emitting the $\gamma$-rays.

@@

Nevertheless, morphological studies of sources in the $\gev$ energy
range using the Fermi LAT is challenging because of the significantly
energy dependent PSF and because of systematic errors in the galactic
diffuse emission. The Fermi PSF varies from over $XXXXXXXXXXXXXXX\deg$
at 100 \mev to $XXXXXXXXXXXXXXXX\deg$ at $100 \gev$ so the higher energy
photons are significantly more important for resolving a source. On
the other hand, we tend to run out of photons at higher energies so a
detailed statistical test must be used to maximize our sensitivity.

@@

Performing morphological studies of sources in the $\gev$ energy range using
the Fermi LAT is challenging primarily because of the significantly energy
dependent PSF and because of systematic errors in the
galactic diffuse emission. The Fermi PSF varies from over $XXXXXXXXXXXXXXXX\deg$
at 100 \mev to $XXXXXXXXXXXXXXXX\deg$ at $100 \gev$ and so the higher energy photons are
significantly more important for studying the shape of sources. One could
ideally look at just the highest energy photons, but the trade off is that
for typical sources many more source photons are expected at low energy.
For this reason, it is especially important to use as much of the data
as possible.

@@

\pointlike additionaly bins the Fermi $\gamma$-rays by their conversion
type, which corresponds to whether the photon converted in the front or
back part of the detector. Since the PSF is smaller for front entering
photons, one would expect binning each conversion type independently to
improve our sensitivity.

@@

Then, the extended source's position error is calculated using the
same code employed to calculate position errors of point sources. This
code is used to calculate the localization error of all point sources
kn the catalog\cite{second_cat}. It works by fitting an ellipse to the
likelihood surface near the maximum and using this fit ellipse to estimate
an elliptical localization errors. By reporting two localization errors
as well as a localization error angle, this method explicitly takes care
of covariance between the two position coordinates.

% Make an extension profile plot.
% show it in figures and describe it.

@@

This is demonstrated in figure~\ref{compare_disk_gauss}.  The plot shows
(in black) the PSF that would be observed for a source of spectral
index 2. The dashed red (blue) line shows the probabability density
function for a Gaussian (uniform disk) of width $0.5\deg$.  The solid
red (blue) line shows the probability density function for the same
Gaussian (uniform disk) convolved with the PSF of an index 2 source.
figure~\ref{compare_disk_gauss} shows the analysis using an energy range
betewen 1 \gev and 100 \gev whereas the right plot shows an analysis using an
energy range from 10 \gev to 100 \gev.

@@

Finally, we note that comparing observed morphologies to 
counts maps can be misleading.  This effect can best be seen
in figure~\ref{compare_r68} where we have simulted a
extended source with 
a spatial model that is a uniform disk of extension 0.5\deg,
integral flux between 1 \gev and 100 \gev of $3e-8\ph/\cm^2/s$,
and a spectral index of 2.

@@

We validated \pointlike for extended source
analysis using a series of Monte Carlo studies.

@@

The half $\chi^2_1$ distribution was found by a Monte Carlo study
to hold for the special case of source finding of EGRET sources using the
maximum likelihood technique.  They were comparing the hypothesis of a
source to the hypothesis of just a background\cite{mattox_egret}.
Their argument for why this null hypothesis distribution was reasonable
was that half of the time, statistical underfluctuations would cause
there to be fewer fewer photons then expected where the sources is being
tested. Since the maximim likelihood machinery cannot fit negative
sources (which would lead to negative model predicted for which the
poisson likelihood is undefined), half of the tests would not have the
likelihood improved by adding a source to the model and $\ts=0$.
It is for the case that the null distribution has

@@

We performed a dedicated Monte Carlo study to validate \pointlike for
studying spatially extended sources and also to demonstrate that we can
use $\tsext$ as a test of the statistical significance
of a source's extension.  The Monte Carlo study involved simulating point
sources of varying spectral models on top of a simulated background. The
simulated point source was then fit using \pointlike as both an extended
source and a point source and $\tsext$ was calculated.

Fermi LAT data was simulated using program \gtobssim which is
distributed publically by the Fermi Science Support Center
(\cite{fssc}).  The sources were simulated with a powerlaw spectral
model with 6 Were given 6 $>100$ \mev fluxes ranging from $3\times 10^{-9}
\ph/\cm^2/s$ to $10^{-6} \ph/\cm^2/s$.  The sources were then simulated
with spectral index ($\gamma$) values of 1.5, 2, 2.5, and 3.  These
values were picked to probe a representative range measurable values.
For particular spectral indices, only the fluxes which would create a
statistically significant source detection ($\ts>25$) where selected,
since we are only interested in testing for extension sources with a
significant extent. The sources were simulated on top of an isotropic
powerlaw background with a Sreekumar-like spectrum ($>100$ \mev flux
of 1.5e-5 and spectral index 2.1) (\cite{sreekumar_isotropic}).

The Monte Carlo simulation was perforormed over a one year time interval using
a default rocking profile and an assumed a represntative
livetime fraction of 0.8.
The reconstruction was performed with photons from an energy range of
1 \gev to 100 \gev using 4 energy bins per decade and using the Pass 7 Version 6
source instrument response function (IRFs) to be consistent with the following analysis.

@@

The results of the Monte Carlo simulation are shown in
figure~\ref{ts_ext_mc}.  The cumulative density function (CDF) is plotted
against $\tsext$ For reference, the CDF for $\chi^2_1/2$,
suggested by Wilks' Theorem, is also plotted for comparison. Overall,
there is good agreement between Wilks' theorem, although the agreement is
not perfect.  It should be noted that the discrepancy seems to be worst
for bright sources which seems to imply that numerical errors in the
way \pointlike models the PSF and performs the convolution become more
apparent for sufficiently bright sources.  Other possible reasons for
departure from Wilks' theorem might include \pointlike ignoring energy
dispersion. Energy dispersion will change the shape as a function of
energy. \pointlike may also run into possibly problems with the fitting
algorithm getting stuck in false minimum. It is important to emphasize
that if
our emperical curve lies to the left of the curve we assume to estimate
significance, we would underestimate the statistical significance of
our result. Since this is the case for most of our simluations, we are
reasonable confident that $\sqrt{\tsext}$ can be safely
used as a measure of the statistical significance of a source's extension.

@@

We performed another Monte Carlo study to determine the Fermi LAT's
sensitivity to spatially extended sources. The sensitivity to extension
is defined as the flux at which 
the average value of $\tsext$ for a source is 
e equal to 25 (for a given spectral model). This correseoncs
to an average extension significance of $5\sigma$.  This is not the
sensitivity to detect an extended source but instead the sensitivity to
resolve an already detected source as being extended.

@@

To calculate the LAT's sensitivity to spatially extended sources,
we simulated sources with a uniform radially symmetric
disk morphology on top of a powerlaw Sreekumar-like isotropic
spectrum (\cite{sreekumar_isotropic}).  The Monte Carlo
simulation was perforormed over a one year time interval using a default
rocking profile and an assumed a represntative livetime fraction of 0.8.
For a given extension and spectral index, we calculated the
sensitivity by varying the flux of a source until we found 
when $\tsext=25$. In particular, we hand picked a flux range
where $\tsext<25$ and $\tsext>25$ and simulated 10 points with
a log uniform distribution in this range and performed an extension
test for each of these fluxes. Since the measured signifciance is
a statistical which will vary from realization to realization,
we then repeated this procedure $>15$ times for each spectral
and spatial model.
By fitting a line to the flux vs $\tsext$, we were able to
estimage the flux at which $\tsext=25$.

Our first result is the extension sensitivity for sources of varying
spectral indices. The result is shown in figure~\ref{index_sensitivity}.
We computed the sensitivty for sources with a spectral index of 1.5,
2, 2.5, and 3 and for sources of 20 extensions varying from $0.1\deg$
to $2.0\deg$. The $y$ axis is the 100 \mev to 100 \gev flux of the
source and the $x$ axis is the Monte Carlo extension. The curves then
represent the division in this extension and flux space where one gets
on average $\tsext=25$.
The solid lines of the plot are the sensitivity computed when using
with all photons with energry between 100 \mev and 100 \gev. The dashed lines is the
same sensitivity but only using photons with energy from 1 \gev to 100 \gev.

This results shows that our extension sensitivity is
significantly better for harder sources where there are more high
energy photons. Furthermore, our sensitivity is best for sources with an
extension around $0.5\deg$.  our sensitivity gets significantly worse
for very small sources and gets slightly worse for very large sources.

Futheremore, we note that except for the hardest or largest sources,
our sensitivity is almost entirly whether or not we use photons 
with energy from
100 \mev to 1 \gev. On the other hand, other systematic errors become
increasingly important at low energy so for our extension search we will
not use the photons in this energy range.

We performed a second computation to see how the LAT's extension
sensitivity varied with increasing background, as is at lower latitude
in the galactic plane. To do so, we computed the sensitivity using the
extragalactic barground as well as 10 and 100 times the extragalactic
background. The latter result approximatly represents the intensity of
the galactic diffuse emission near the galactic center. The results are
shown in figure~\ref{diff_factor_sensitivity}. Two plots are presented,
one fitting only in the energy range between 1 \gev and 100 \gev and
one fitting only in the energy range between 10 \gev and 100 \gev. For
each plot, the $y$ axis quotes the flux in the fit energy range and the
simulation was done only for a representivite spectral index 2 source.

@@

The LAT two year catalog presented a search for all
statistically significant sources that are detected in the \gev
sky\cite{second_cat}.  The catalog group's source finding method assumes
that the emission originated as point sources and what was presented was
a list of all sources fit assuming a point source spatial hypothesis. 
Published extended sources were modeled by the catalog group as 
being extended, but no attempt was made to test if other sources in the catalog
could be spatially resolved.

@@

As was shown, we gain very little in sensitivity using photons with energy
below 1 \gev. On the other hand, the large PSF at low energy makes us
more succeptable to systmetic errors arising from source confusion due
to multiple point soruces and succeptable to incorrect modeling of the
galactic diffuse emission.  For that reason, we performed our search
using photons only between 1 \gev and 100 \gev.

We also performed a search for extended sources using photons only between
10 \gev and 100 \gev. Even though we are testing the same sources, this
approach is complimentary because there are regions in the galactic
plane where at lower energies the diffuse emission dominates and is
likely not perfectly modeled. On the other hand, at the highest energies
accessible by the LAT, the galactic diffuse emission will be smaller and
we might expect to detect significantly harder sources. This approach is
also beneficial because harder sources with nearby neighbors that are
softer may end up being incorrectly fit at lower energies but correctly
found looking only at the high energy.  This is also beneficial in the
galactic plane looking for possible extended sources near pulsars which
are very bright at lower energy but cutoff before 10 \gev.  This approach
of using only photons with energy above 10 \gev was succesfully used to detecting
HESS J1825-137 with the Fermi LAT~\cite{fermi_hess_j1825}.

For each catalog source, we modified the spectral index of the source
to be a powerlaw. We did this for two reasons. The first is that for a
source to be significantly extended, we would expect it to be reasonably
hard and therefore be fit reasonably well by a powerlaw.  The seond reason
is that since we begin our analysis at either 1 \gev or 10 \gev, we didn't
expect there to be significant curvature over this smaller energy range.
This method has the advantange that for the following analysis, we can
simply quote the fit powerlaw spectral index.

Finally, when analyzing our region, we automatically removed all other
catalog sources which were within 0.5\deg. This was done because of
a concern that truly extended sources might end up in the catalog as
multiple point sources. These spurious nearby catalogs sources might then
distort our fit enough that it no longer would show up as being extended.
Nevertheless, we investiated for each of our extended candidates if
if we seemed to be legitimatly removing real backgorund sources. And
we employed a dual localiation procedure (described above) to test
all of our candidates for source confusion.

For each catalog source, we modified the source to have a uniform radially
symmetric disk spatial shape and used \pointlike to simultaneously fit
the position and extension of the source.  We then selected all of the
sources for which \pointlike found $\tsext>16$.

@@

We expect most extended sources to be inside teh galactic plane
since extragalactic sources are typically too far away to be
resolvable. Unfortunately, the \gev emission in the galactic plane
is dominated by anisotropic galactic diffuse emission.  Modeling the
galactic diffuse emission and finding sources on top of this emission is
difficult~\ref{first_diffuse_paper}. This issue has been discussed in the
Fermi catalog~\ref{first_cat,second_cat}. Furthermore, the galactic plane is
very crowded and so it is offent difficult to get a good model for the
region around a catalog source.  Because of this, finding a source with
$\tsext>16$ is not a sufficient test and we perform several crosschecks
to ensure our overal region is well modeled.

As an example of why this issue, we describe our test of P72Y3047 for
extension.  Figure~\ref{example_bad_fit} shows a smoothed counts map
using photons with energy above 1 \gev. In the map, the galacitc and isotropic diffuse
emission subtracted.  A two year catalog source P72Y3047 was found at the
position $(l,b)=(37.79\deg,1.14\deg)$.  By looking at the smoothed map,
we see that there appears to be overall some large scale residaul in the
diffuse emission along the galactic plane. Furthermore, it appears that
we don't have a great model of the nearby catalog sources. As a result,
P72Y3047 fits to a very large extension of 1.37\deg and the result is
statistically significant with \tsext=51.6.  By looking at the residuals,
it is clear that this complicated region is not well fit even though
the source's extension is statistically significant. For this reason,
we have to perform a series of followup analysis for all of the soruces
that are found with \tsext=16.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml

First, we took the spatial and spectral model found using \pointlike
and performed a crosscheck by refitting the spectral parameters using
\gtlike.  We used a square ROI $14.1\deg$ in diamater with $1/16\deg$
pixels and the same energy binning. We found that $1/16\deg$ bins were
fine enough for \gtlike to get an extension significances consistent
with \pointlike.  Using \gtlike, we found a second measure of \tsext and
we only considered a source to be significanlty extended if \pointlike
and \gtlike both agreed.  For the new extended source candidates, with
table~\ref{alt_diff_model_results} shows a comparison of their the value
of \tsext found by \gtlike and \pointlike.  Everywhere else, the values
from \gtlike are used.

Additionaly, we generated for each extended source candidate residual
test statistic (\ts) maps.  The residual \ts map is a map of the \ts
value evalualted at each point by adding onto the map at that position
a new source of spectral index 2 and finding the increase in lieklihood
when fitting its flux. \ts maps are especially useful as looking for
new sources or other residual featuers that are not included in the
model. They also have the advantagous feature that the aboslute scale
can be used to clearly look for when the residual is signifciant.
The downside is they are biased to look for residual features
that look like the PSF and so are not as sensitive to diffuse residual
features. Furthermore, since by defintion \ts>0, \ts maps can only show
positive fluctuations. Nevertheless, \ts maps were useful for
assessing the overal quality of the fit. Figure~\ref{res_tsmap}
shows an exmaple residual \ts map for the extended source IC443.

For each candidate, we also generated counts slices. A counts slice
plots the sum of all counts in the longitudinal direction within 2\deg in
the latitudinal direction.  The same plot is also shown integrating in
the longitudinal direction.  On top of the plot is overlayed the model
predicted counts assuming the source has a disk spatial model, assuming
the source is a point source, and fitting just the diffuse emission in
the region.  An example counts slice is shows for the extended source
IC443 in figure~\ref{counts_slice}. Counts slices were especially useful
for looking at how well the overal diffuse emisison model fits the region
near the sources.

For each candidate, we also made was a radial integral of the data. This
plot shows the sum of all counts within a given distance from the
source's center. The plot is binned uniformly in $\Delta \theta^2$ so
that an isotropic background will be flat in the plot. On top of this
plot is overlayed the model predicted counts assuming the source has a
disk spatial model, assumgng the source is a point source, and fitting
just the diffuse emisison in the region. An example radial integral
is shown for the extended source IC443 in figure~\ref{radial_profile}.
This plot is best for comparing the observed counts to the PSF.

Finally, for each source, we performed the dual localization method
described in section~\ref{dual_localization_method}.  We 
did so by replacing the source with two sources each containing
half of the original flux and then using \pointlike to
simulatnoulsy fit the postion and spectrum of these sources.
We then used \gtlike to refit the spectrum of these
sources using the best fit positions. For the
new extended sources found, we summarize the results of
this dual localization test and compare the values of \tsext
to \tsinc in table~\ref{dual_localization_results}.

for sources that seemed promising, we performed in a case
by case basis a refit of the region where we would move or delete
background catalog sources until it was clear that we had obtained on
a case by case a good overal fit of the region. This was necessary in
many situations whrere the position of background sources in the catalog
was not consistent and when the offet of the background sources biased
the fit extension of our source. This was likey the case because the
catalog positions of the backgorund sources came usign all photons above
100 \mev and the low energy localization disagreed with the higher energy
localization we found. This also happened in certain situations because
the localization of a backgorund source dependend upon correctly modeling
the extended source as being extended and so once we fit our source for
extension, we had to relcoalizae the backgorund source.  For each source,
we descirbe in a case by case basis any modifications which were required
to obtain a good fit of our source and the surrounding region.

@@

We validated our method by testing for extension all of the sources
which the Fermi Collaboration had previously published on.  
This list includes the extended sources where were modeled
in the 2FGL catalog\cite{second_cat}.
This list contains the Large Magelenic Cloud and the Small Magilenic Cloud
which were first published in \cite{LMC,SMC}.
We also included the Supernova rememants IC443, W28, W30,
W44, W51C, and the Cygnus Loop which were 
first published in \cite{ic443,w28,w30,w44,w51c,cygnus_loop}.
We include the Pulsar Wind Nebulae Vela X, MSH 15-52, and 
HESS J1825-137\cite{velax,msh1552,fermi_hess_j1825}.
In addition to those sources included by the catalog, we include in
this list the supernova rememnat RX J1713.7-3946 which was recently
published (\cite{RXJ1713}).
